---
title: "joanna_post1"
description: |
  Description here
author:
  - name: Joanna Wright
    url: https://example.com/norajones
date: 05-08-2020
output:
  distill::distill_article:
    self_contained: false
---

This post describes examples of using iteration, parallel iteration, and writing functions in order to loop through a data set to extract information and create plots.

## Set-up

```{r setup }
knitr::opts_chunk$set(echo = TRUE)

```

<aside>
Libraries for this post
</aside>
```{r}
library(ggplot2)
library(tidyverse)
library(janitor)
library(purrr)
```


Loading the data. For details, see [Brock's post](link here).

```{r loading the data, results='hide'}

files <- list.files(here::here("data"),
                    full.names = TRUE)
files
d <- read_csv(files[1]) %>%
  clean_names()

```

## Task 1: Plotting mobility trends by country 

Let's say we want to create a separate plot for mobility trends in each country. For this example, let's start with one category - workplaces. First, we need to `group_by` country and `nest()` the rest of the data. This creates a list of tibbles, one for each country. I've called `head()` here so that we can see what the first few rows look like. 

```{r}
d %>%
  group_by(country) %>% 
  nest() %>%
  head()
```

Now we can use this as the input for our plots. `pmap()` is a variation of `map()` which allows us to apply a function to multiple vectors simultaneously.

```{r parallel iteration, workplaces mobility trends for each country}

plots_workplaces <- d %>%
  group_by(country) %>% 
  nest() %>% 
  mutate(plots = pmap(list(country, data),  
                      ~ggplot(..2, aes(date, workplaces)) +
                        geom_point() + 
                        geom_line()))

# test to see what a couple look like:
plots_workplaces$plots[1]
plots_workplaces$plots[2]

# or we could see all of them: 
# plots_workplaces$plots[1:19]
```

To add a unique title to each plot, use `{..1}` to select the 1st column, country, to insert country names into the title. 

```{r adding unique title to each plot with country name}

# adding unique title: 
plots_workplaces1 <- d %>%
  group_by(country) %>% 
    nest() %>%
    mutate(plots = pmap(list(country, data),
    ~ggplot(..2, aes(date, workplaces)) +
      geom_point() + 
      geom_line() + 
      labs(title = paste("Mobility trends for workplaces in {..1}"),
           x = "Date", 
           y = "% change")))

# testing: 
plots_workplaces1$plots[1]
plots_workplaces1$plots[2]

# or we could see all of them: 
# plots_workplaces1$plots[1:19]
```

So far, we've looked at mobility trends in one particular category (workplaces), for each country. 
Now let's generalize this with a function that we could apply to any selection of mobility categories (rather than just workplaces). 

Our function will have 2 arguments: the data frame, and the mobility category. The x-axis variable, time, will remain the same, so it is supplied in the body of the function rather than as an argument. 

```{r function for creating plots }

# assign nested data as object: 
by_country <- d %>%
  group_by(country) %>% 
    nest()

# First let's get the plot for just one country (the first one, as selected using by_country$data[[1]]), and one category, parks. 

ggplot(by_country$data[[1]], aes(date, parks)) +
      geom_point() + 
      geom_line() + 
      labs(title = paste("Overal mobility trends in parks"),
           x = "Date", 
           y = "% change")

# Now generalize to a function: 
create_plot <- function(df, outcome) {
  call <- as.character(match.call())
  ggplot(df, aes(date, {{outcome}})) +
      geom_point() + 
      geom_line() + 
      labs(title = paste("Overall mobility trends in", call[3] ), 
           x = "Date", 
           y = "% change")
}

create_plot(by_country$data[[1]], parks)

# loop create_plot through the by_country data frame

# map(by_country, ~create_plot(df = by_country, outcome = .x))
# PROBLEM HERE: WHY? 
# Error: Aesthetics must be valid data columns. Problematic aesthetic(s): x = date. Did you mistype the name of a data column or forget to add after_stat()?

# once we figure the above out, could assign to plot_list and use walk() to print all plots:  
# walk(plot_list, print)

```

```{r function for creating plots, continued}

# Alternative approach: we could restructure the data to have column indicating categories, then nest by country AND category: 

by_country_mobil_cat <- d %>%
  pivot_longer(
    cols = 4:9,
    names_to = "mobil_category",
    values_to = "perc_change"
  ) %>%
  nest(-country, -mobil_category)

# Then loop through data and apply the plotting function we created: 

# by_country_mobil_cat %>%
# pmap(list(country, mobil_category, data), ~create_plot(df = by_country_mobil_cat, outcome = .x))

# THIS IS CLOSE? But why does the error message say "object 'country' not found"?? 

```


## Task 2: How do infection rates relate to mobility trends, by country?

Fitting a linear model to see how mobility changed as infection rate changed in each country: 

```{r results="hide"}

d1 <- d %>%
  mutate(total_mob = rowSums(.[4:9]))

lm_by_country_lists <- d1 %>% 
  split(.$country) %>% # split into separate tbl for each country
  map(~lm(total_mob ~ total_cases, data = .)) # for each, fit lm
# Outputs a nested list

```

Extracting slopes: 

```{r }
slopes <- lm_by_country_lists %>%
    map(coef) %>%
    map_dbl(2)
slopes
```

## Things to work on:

Resolve bugs in 'function for creating plots' code chunks.

Improve plot visualization 

Repeated use of code? 
