---
title: "Rowley_Post"
description: |
  We are getting started with setting up a project, working directory, subfolders, and loading data.
author:
  - name: Brock Rowley
    url: https://example.com/norajones
date: 05-11-2020
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)
library(usethis)
```

# Set up a working directory complete with an Rproject and subfolders. Load data with a function.

## Let's get set up!
We want to keep all our data, analyses, and text in one, self-contained, single folder, called the working directory. RStudio offers a helpful “Projects” interface. Let’s go through the steps for creating an “R Project”.

Start RStudio.
Under the File menu, click on New Project. Choose New Directory, then New Project.
Enter a name for this new folder (or “directory”), and choose a convenient location for it. This will be your working directory for the rest of the day (e.g., ~ /funprog_blog).
Click on Create Project.

Use the here() function so the results in the path are always relative to our projects root directory.

# Create a new package -------------------------------------------------
path <- file.path(tempdir(), "mypkg")
create_package(path)

## Data from a CSV
The most common way to get data into R is to store it as a CSV.

base R has a read_csv() command from readr. readr is part of the tidyverse, and read_csv() is fast.

```{r}
files <- list.files(here::here("data"),
                    full.names = TRUE)

d <- read_csv(files[1]) %>%
  janitor::clean_names()
```

You can also read a CSV file straight from the Internet via a URL. CSVs should have the file extension .csv.

```{r load_url}
url <- "https://www.kaggle.com/gustavomodelli/covid-community-measures"
cool_name <- read_csv(url)
View(cool_name)
```

To write a data frame to a CSV, use the write_csv() function. By default, the file will be written to the working directory.
```{r}
write_csv(data, path = "cool_name.csv")
```

