---
title: "GettingStarted_Post"
description: |
  We are getting started with setting up a project, working directory, subfolders, and loading data.
author:
  - name: Brock Rowley
    url: https://example.com/norajones
date: 05-11-2020
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)
library(usethis)
```

## We will set up a working directory, complete with an Rproject and subfolders. We will show you how to load your data and save your data.

# The typical way to get set up:
You are probibly framilar with this process as a typical way to start any RStudio project. The process goes something like this:  
Start RStudio.  
Under the File menu, click on New Project.  
Choose New Directory, then New Project.  
Enter a name for this new folder (or “directory”), and choose a convenient location for it. This will be your working directory for the project (e.g., ~ /funprog_blog).  
Click on Create Project.

# Using the here function to help with file pathways

Let's start this time by figuring out one of the most difficult aspects of learning to code initially... where are we?? OR... where does my computer think we are pointed? file pathways are difficult to navigate, keep straight and understand when you first begin to code. We will use the here() function so the results in our path are always relative to our projects root directory. First understand where we are.

```{r directory_locate_here, include = TRUE, echo = TRUE}
here::here()
# create_setup(here::here())
# create_setup(here::here("temporary-wd"))
```

Now that we know the file pathway, this thing: "[1] "/Users/brockrowley/Phd in SPED/2020Spring/funprog_blog" we know where we are (root).

#Thuy: Hi, I'm not sure, but I guess it's a way to link the result of the pathway I run in line 36, and show it down here. I mean I should see: "Now that we know the file pathway, this thing: "[1] "/Users/thuyn/Desktop/EDLD610 data science/funprog_blog", OR anyone who run your code should see their path, instead of your directory, which is "[1] "/Users/brockrowley/Phd in SPED/2020Spring/funprog_blog"


# Create a new directory here

We want to keep all our data, analyses, and text in one, self-contained, single folder, called the working directory. We will use the fs package to do this work for us.

```{r setting_up_directory, include = TRUE, echo = TRUE}
# Create a (temporary in blog) directory with ecco = FALSE for blog but, ecco = TRUE for output to blog
fs::dir_create(here::here("temporary-wd"))

# Write a function that (basically) does this below: 
usethis::proj_activate(here::here("temporary-wd"))

create_setup <- function(dir = getwd()) {
  to_create <- c("data", "scripts", "plots")
  purrr::walk(to_create, ~{
    fs::dir_create(file.path(dir, .x))
  })


}

create_setup(here::here("temporary-wd"))
```

# Thuy: when I run this code, a new Rstudio session pops up. You may want to comment out line 65 and give the audience a heads up (like what you already prepared below) --> then learners can de-comment it and run :) 

In the code chunk above, we used "temporary-wd" as a placeholder for whatever you want to call your project. If you replace "temporary-wd" with whatever you want to call your project, you will create a new RStudio project, complete with three new folders inside (data, scripts, plots) to get started.

# Data from a CSV
The most common way to get data into R is to store it as a CSV. Drop your data file into the folder we created above called "data" and run the following code chunk to load your data.
base R has a read_csv() command from readr. readr is part of the tidyverse, and read_csv() is fast.

```{r load_data, include = TRUE, echo = TRUE}
files <- list.files(here::here("data"),
                    full.names = TRUE)

d <- read_csv(files[1]) %>%
  janitor::clean_names()
```

# Data from a URL

You can also read a CSV file straight from the Internet via a URL. CSVs should have the file extension .csv.

```{r load_url, include = TRUE, echo = TRUE}
#url <- "https://www.putyourcoolurl/linktofile/here/download.csv"
#covid_data <- read_csv(url)
#View(covid_data)
```

# Saving your data to folder

To write a data frame to a CSV, use the write_csv() function. By default, the file will be written to the working directory.

```{r write_data, include = TRUE, echo = TRUE}
#write_csv(data, path = "cool_name.csv")
```

The following code chunk is here to delete the temporary working directory (temporary-wd) used to show how the function works.

```{r last-chunk, echo = FALSE}
fs::dir_delete(here::here("temporary-wd"))
```