---
title: "How to Subset Data and Output Basic Descriptives"
description: |
  A tutorial on writing a function to group your data and generate basic descriptives. 
author:
  - name: Stephanie Gluck & Akhila Nekkanti
    url: https://s-gluck.github.io/funprog_blog
date: 05-23-2020
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 2
---

```{r setup, include= FALSE}

library(tidyverse)
library(lubridate)
library(glue)
library(janitor)
library(purrr)
library(reactable)

```

# Project Notes
### This section will be deleted or moved for final version

To create a new page on your blog 
`distill::create_post("post_title")`

To tackle the outline: 

1) Set up a working directory complete with an Rproject and subfolders - Brock
 
2) Subsetting data and outputting basic descriptives - Akhila & Stephanie 
  - Custom functions for means and standard deviation
  - Rename Akhila's first post - Dataset Descriptives Tutorial
  - Idea: Subset by week (function to separate dates by incrememnts of 7), location, and country, to get average activity per week in each location.

3) Loop through regression/other stats models to extract plots - Joanna

4) Creating folders to export plots with meaningful paramters (e.g., names, plot dimensions) - Akhila & Stephanie 

Data:
https://www.kaggle.com/gustavomodelli/covid-community-measures

More info about the data: 
https://support.google.com/covid19-mobility/answer/9824897?hl=en&ref_topic=9822927

**Class Requirement**
No code is used repetitively (no more than twice) [10 points]

More than one variant of purrr::map is used [10 points]

At least one {purrr} function outside the basic map family (walk_*, reduce, modify_*, etc.) [10 points]

At least one instance of parallel iteration (e.g., map2_*, pmap_*) [10 points]

At least one use case of purrr::nest %>% mutate() [10 points]

At least two custom functions [20 points; 10 points each]
   - Each function must do exactly one thing
   - The functions may replicate the behavior of a base function - as noted above this is about practicing the skills you learn in class



# Introduction 

For the tutorial, we will use the [COVID Community Measures](https://www.kaggle.com/gustavomodelli/covid-community-measures) dataset from Kaggle. We hope that you will be able to generalize the function we covered to be applicable to your own dataset. The current dataset contains Community Mobility Reports describing community activities in a variety of settings from Google and the incidence of COVID-19 diagnosis and fatalities from John Hopkins CSSE. 

Community activity is defined as the change in activity from the baseline days. A baseline day is the median value from the 5‑week period between January 3rd to February 6th, 2020. 


## Loading the Data 

Here are the packages used in our tutorial: 

```{r, echo = TRUE}

#list the packages that readers need for tutorial when finish with post 
#consider specifying which tools from each package we will need. 
library(tidyverse) 
library(janitor)
library(lubridate)
library(purrr)

```

Our blog also includes a post on how to automate your data loading. [You can check out the post here](insert link to Brock's Post)

Here is how the data looks like: 


```{r, echo = FALSE}

files <- list.files(here::here("data"),
                    full.names = TRUE)

d <- read_csv(files[1]) %>%
  clean_names()

reactable(d,  
          height = 500, 
          striped = TRUE, 
          compact = TRUE, 
          outlined = TRUE)

```


The data contains the following variables: 

```{r variable_names, echo = TRUE}

names(d)

```

from the following countries: 

```{r countries, echo = TRUE }
#base R method
#table(d$country)

#using tidyverse
d %>% count(country, sort = TRUE)
```

# Function Tutorial 

## Subsetting Data 


```{r}
#add day of week column and a week category column in order to split by 7 days interval. Label = TRUE will print actual day of the week, e.g., Sunday (1)

d <- d %>% 
  mutate(week = cut(date, "week", start.on.monday = FALSE, labels = FALSE), 
         day_of_week = wday(date, label = TRUE))

```




```{r, eval = FALSE, echo = TRUE}

#Stephanie comment: MAJOR work in progress!!!! I eval = F and echo = T for now, need to fix later

str(d)

d_split <- split(d, d$country)

d_split[[1]]$country

d_split2 <- d %>%
    nest(-country)
d_split2 

str(d_split2$data)

map_dbl(d_split2$data, nrow)

map_dbl(d_split2$data, ncol)

map_dbl(d_split2$data, ~mean(.x$residential))


map(d_split, ~map_chr(.x$country, ))


pokemon <- map(pokes, ~map_chr(.x$pokemon, ~.x$pokemon$name))

map_chr(d_split, d_split[[.x]]$country)

map(d_split, list("country")) %>% 
  map_chr(country)

country_names <- map_chr(d_split, ~(.x$country))
country_names


ability <- map_chr(abs, "name")

#extract length for 1 list
length(abs[[1]]$pokemon)

#convert into map 
count <- map(abs, list("pokemon")) %>% 
  map_dbl(~length(.x))

#another way to coutn # of pokemon with each ability: 
map_dbl(abs, ~length(.x$pokemon))

#create tibble 
tibble(Ability = str_to_title(ability), 
              `Number of Pokemon` = count)

```

## Function for General Descriptives

```{r, eval = FALSE, echo = TRUE}

#function from lab 4 incase is helpful for our tutorial
#I eval = F and echo = T for now, need to fix later

##Latest version of function 5/25 - akhila


descriptives <- function(df, outcome, ..., .fun = list(n = function(x) length(x),
                                             n_valid = function(x) sum(!is.na(x)),
                                             n_missing = function(x) sum(is.na(x)),
                                             mean = function(x) mean(x, na.rm = TRUE),
                                             sd = function(x) sd(x, na.rm = TRUE),
                                             min = function(x) min(x, na.rm = TRUE),
                                             max = function(x) max(x, na.rm = TRUE))) {
  library(dplyr)
  df_name <- deparse(substitute(df))
  outcome_name <- deparse(substitute(outcome))
            if(!is.data.frame(df)) { 
              
               stop(df_name, " is not a data frame")
            } 
  
            if(!is.numeric(pull(df, {{outcome}}))) {  ## only works with {{ }}
              
               stop(outcome_name, " is not numeric")
            }
  
            if(is.numeric(pull(df, ...))) {
              
              warning(" Grouping variables are not categorical")
            }
 
  df %>% 
    group_by(...) %>%  ##doesn't work with {{ }}
    summarise_at(vars({{outcome}}), 
                 .funs = .fun)
}


listof_fun <- list(
  n = function(x) length(x),
  n_valid = function(x) sum(!is.na(x)),
  mean = function(x) mean(x, na.rm = TRUE),
  sd = function(x) sd(x, na.rm = TRUE),
  min = function(x) min(x, na.rm = TRUE),
  max = function(x) max(x, na.rm = TRUE)
)


smry_function <- function(df, outcome, ..., .fun = list(n = function(x) length(x),
                                             n_valid = function(x) sum(!is.na(x)),
                                             mean = function(x) mean(x),
                                             sd = function(x) sd(x),
                                             min = function(x) min(x),
                                             max = function(x) max(x))) {
  df %>% 
    group_by(...) %>% 
    summarise_at(vars({{outcome}}), .funs = .fun)
}

smry_function(iris, Sepal.Width, Species)



smry_function2 <- function(df, outcome, ..., .fun = listof_fun
) {
  library(dplyr)
  df %>% 
    group_by(...) %>% 
    summarise_at(vars({{outcome}}), .funs = .fun)
}

smry_function2(iris, Sepal.Width, Species)

smry_function2(airquality, Ozone, Month)



```

